```{r todaysDate,echo=FALSE}
td <- format(Sys.Date(), format="%A, %B %d, %Y")
```
---
title: "Assignment 1"
author: "Micho Janssen"
date: `r td`
output: html_document
---

<br><br>
<h4>Loading and preprocessing data </h4>

Load data
```{r loadData}
act <- read.csv("activity.csv")
```
Remove rows with missing values
```{r removeNAs}
activity <- act[complete.cases(act),]
```

<br><br>
<h4>Mean total tumber of steps taken per day: </h4>
```{r stepsPerDay}
y <- aggregate(activity$steps,by=list(activity$date),sum)
names(y) <- c("day","totalsteps")
spdmean <- format(mean(y$totalsteps),digits=6)
spdmedian <- median(y$totalsteps)
hist(y$totalsteps,ylab="",xlab="Steps",main="Total Steps per Day")
```

Total Number of Steps Taken per Day:

Mean:     `r spdmean`

Median: `r spdmedian`

<br><br>
<h4>Average daily activity pattern</h4>

```{r averageActivity}
y <- aggregate(activity$steps,by=list(activity$interval),mean)
names(y) <- c("interval","meansteps")
maxInterval <- y[which.max(y[,2]),1]
plot(y$interval,y$meansteps,type="l",ylab="Average steps",xlab="Interval",main="Average Steps per Interval")
```

The 5-minute interval with the maximum number of steps is: `r maxInterval`

<br><br>
<h4>Imputting missing values</h4>

```{r rowsMissingValues}
nmissing <- nrow(act) - nrow(activity)
```
Number of rows with missing values: `r nmissing`

```{r imputtingMissingValues}
# If 'steps' value missing imput average of same interval day before and after. If the new value is still NA, use day before value. At beginning use average of 2 days after, at end use average of 2 days before
numIntervals <- length(act[which(act$date == act[1,2]),2])
actRows <- nrow(act)
for (i in 1:actRows) {
    y <- act[i,1]
    if (is.na(y)) { # Begining of Frame
        if (i <= numIntervals) {
            act[i,1] <- (act[i+numIntervals,1]+act[i+2*numIntervals,1])/2
        } else if ((i+numIntervals) > actRows) { # end of Frame
            act[i,1] <- (act[i-numIntervals,1]+act[i-2*numIntervals,1])/2
        } else { # Take average of day before and day after
            if (is.na(act[i+numIntervals,1])) { # Still NA, use day before values
                act[i,1] <-  act[i-numIntervals,1]
            } else {
                act[i,1] <- (act[i-numIntervals,1]+act[i+numIntervals,1])/2
            } # .. (is.na(act[i+numIntervals,1))
        } # ..(i <= numIntervals)
    } # is.na(
} # .. in 1:actRows
# check any NAs with - act[!complete.cases(act),]

# Values filled in, recalculate mean and median
y <- aggregate(act$steps,by=list(act$date),sum)
names(y) <- c("day","totalsteps")
spdmean2 <- format(mean(y$totalsteps),digits=6)
spdmedian2 <- format(median(y$totalsteps),digits=5)
hist(y$totalsteps,ylab="",xlab="Steps",main="Total Steps per Day")
```

**Total Number of Steps Taken per Day:**

|         | Imputing missing Values  | Removing Missing Values |
|---------|--------------------------|-------------------------|
|  Mean   |    `r spdmean2`          |    `r spdmean`          |
|  Median |    `r spdmedian2`        |    `r spdmedian`        |

<br><br>
<h4>Weekday and Weekend Activity </h4>

```{r weekdaysWeekends}
y <- as.POSIXlt(act[,2])$wday
y <- ifelse(y>0 & y<6,1,0)  # 1=weekdays, 0=weekends
act$wde <- y                # add factor(y) to act df
y <- aggregate(act$steps,by=list(act$interval,act$wde),mean)
names(y)=c("interval","wday","averagesteps")
y$wday <- factor(y$wday, labels = c("Weekends","Weekdays"))
library(ggplot2)
qplot(interval,averagesteps,data=y,geom="line",xlab="Interval",ylab="Average",facets=wday~.)
```
